---
title: "STAT-GB Forecasting Timeseries - citibike-realtime-analysis"
author: "Amelia Chu"
date: "4/10/2018"
output:
  html_document: default
  pdf_document: default
---

```{r libraries, echo=FALSE}
library("forecast")
library("tseries")
```

Loading in Data
```{r data}
data <- read.csv("/Users/chuamelia/Google Drive/Forecasting Time Series/citi-bike/ts-realtime-analysis-bike/status_161.csv")
date <- as.POSIXlt(data$last_updated)
time <- 1:length(date)
status <- data$num_bikes_available + 1 
```

Taking logs and differences
```{r}
log.status <- log(status)
diff.log.status <- c(NA, diff(log.status))
diff2.log.status <- c(NA, diff(diff.log.status))
```

### Figure 2. Plotting timeseries of Bikes Available, its log and differenced logs.
```{r fig-2, fig.width=12, fig.height=8}
par(mfrow=c(2,2))
plot(date, status, type="l",
     xlab="Date", ylab="Bikes Available")
plot(date, log.status, type="l",
     xlab="Date", ylab="Log Bikes Available")
plot(date, diff.log.status, type="l",
     xlab="Date", ylab="Differenced Log Bikes Available")
plot(date, diff2.log.status, type="l",
     xlab="Date", ylab="Differenced2 Log Bikes Available")
```

### Figures 2b, 3a, 3b. Timeseries of Log Bikes Available and ACF/PACF
```{r fig.width=12, fig.height=8}
# Time series plot
layout(matrix(c(1,1,2,3), 2, 2, byrow = TRUE))
plot(date, log.status, type="l",
     xlab="Date", ylab="Log Bikes Available")

# ACF and PACF

Acf(log.status, na.action = na.pass)
log.status.acf <- Acf(log.status, na.action = na.pass, plot = FALSE)
pacf(log.status, na.action = na.pass)
log.status.acf.table <- data.frame(log.status.acf$lag,  log.status.acf$acf)[-1,]
```

### Figures 2c, 3c, 3d. Timeseries of Differenced Log Bikes Available and ACF/PACF
```{r fig.width=12, fig.height=8}
layout(matrix(c(1,1,2,3), 2, 2, byrow = TRUE))
# Time series plot
plot(date, diff.log.status, type="l",
     xlab="Date", ylab="Differenced Log Bikes Available")
# ACF and PACF
Acf(diff.log.status, na.action = na.pass)
pacf(diff.log.status, na.action = na.pass)
```

### Figures 2d, 3e, 3f. Timeseries of Twice-Differenced Log Bikes Available and ACF/PACF
```{r fig.width=12, fig.height=8}
layout(matrix(c(1,1,2,3), 2, 2, byrow = TRUE))
# Time series plot
plot(date, diff2.log.status, type="l",
     xlab="Date", ylab="Differenced2 Log Bikes Available")
# ACF and PACF
Acf(diff2.log.status, na.action = na.pass)
pacf(diff2.log.status, na.action = na.pass)
```

### Figure 4. Computing the AICc values for ARIMA Candidate Models
```{r }
d <- 1

# choose p, q with AICc
for (include.constant in c(FALSE, TRUE)) {
    for (p in 0:4) {
        for (q in 0:4) {
            # work-around bug in R by manually differencing
            fit <- Arima(diff(log.status), c(p,0,q),
                         include.constant=include.constant, method="ML")
             cat("ARIMA",
                "(", p, ",", d, ",", q, ")",
                "(constant=", include.constant, ")",
                " : ", fit$aicc, "\n", sep="")
             #cat( p, ":", d, ":", q, ":",
            #     ":", include.constant, "",
             #    " : ", fit$aicc, "\n", sep="")
        }
    }
}

```

### Figure 5. Final Estimates of ARIMA Parameters

Here is code to fit the ARIMA model, then compute residuals and the fitted values:

```{r}
fit.mean <- Arima(log.status, c(3, 1, 3), include.constant=FALSE)
summary(fit.mean)
```

## Figure 6. Modified Box Pierce (Ljung-Box) Diagnostics
```{r}
Box.test(diff.log.status, lag = 24, type = c("Box-Pierce", "Ljung-Box"), fitdf = 0)
Box.test(diff.log.status, lag = 36, type = c("Box-Pierce", "Ljung-Box"), fitdf = 0)
Box.test(diff.log.status, lag = 48, type = c("Box-Pierce", "Ljung-Box"), fitdf = 0)
```
### Figure 7. ARIMA Residual Plots

Here are the residuals, with the last 10 residuals printed out:
```{r}
resid <- residuals(fit.mean)
tail(resid, n=10)
```


Here are the fitted values, with the last 10 fitted values printed out:
```{r}
f <- fitted.values(fit.mean)
tail(f, n=10)
```

### Figure 10. Forecast Intervals for ARIMA Model
Here is the one step ahead forecast and 95% forecast interval:

```{r}
forecast(fit.mean, h=1, level = 95)
```

Here is a plot of the residuals:

```{r fig-7a}
plot(date, resid, type="l",
     xlab="Date", ylab="Bikes Available Residuals")
```

Here are the ACF and PACF of the residuals:

```{r fig.width=12, fig.height=6}
# ACF and PACF
par(mfrow=c(1,2))
Acf(resid, na.action = na.pass)
Pacf(resid, na.action = na.pass)
```

Here are the ACF and PACF of the squared-residuals:

```{r fig.width=12, fig.height=6}
# Add ACF, PACF of squared residuals.
par(mfrow=c(1,2))
Acf(resid^2, na.action = na.pass)
Pacf(resid^2, na.action = na.pass)
```

### Figure 8. AICc for Candidate ARCH/GARCH Models

Here are the AICc values for the ARCH(q):

```{r}
q <- 0:10
loglik <- rep(NA, length(q))
N <- length(resid)

for (i in 1:length(q)) {
    if (q[i] == 0) {
        loglik[i] <- -0.5 * N * (1 + log(2 * pi * mean(resid^2)))
    } else {
        fit <- garch(resid, c(0,q[i]), trace=FALSE)
        loglik[i] <- logLik(fit)
    }
}

k <- q + 1
aicc <- -2 * loglik  + 2 * k * N / (N - k - 1)

print(data.frame(q, loglik, aicc))

```
Exporting the arch aicc table for website:

```{r}
arch.aicc.table <- data.frame(q, loglik, aicc)
write.csv(arch.aicc.table, file="/Users/chuamelia/Desktop/ameliachu.github.io/realtime_bike/arch-aicc.csv", row.names = FALSE)
```


Here is the AICc for the GARCH(1,1):

```{r}
fit <- garch(resid, c(1,1), trace=FALSE)
loglik <- logLik(fit)
k <- 2
aicc <- -2 * loglik  + 2 * k * N / (N - k - 1)

print(data.frame(loglik, aicc))
```


```{r echo=FALSE}
# Exporting the garch aicc table for website:
garch.aicc.table <- data.frame(loglik, aicc)
write.csv(garch.aicc.table, file="/Users/chuamelia/Desktop/ameliachu.github.io/realtime_bike/garch-aicc.csv", row.names = FALSE)
```

### Figure 9. Model Output for GARCH
Here are the summary and log likelihood of the selected model:

```{r}
fit.var <- garch(resid, c(1,1), trace=FALSE)
summary(fit.var)
logLik(fit.var)
```
### Figure 10. Forecast Intervals for ARIMA-ARCH Model

```{r}
a0 <- coefficients(fit.var)["a0"]
a1 <- coefficients(fit.var)["a1"]
b1 <- coefficients(fit.var)["b1"]
f1 <- 1.943208

# conditional variance:
h1<- a0 + a1*(tail(fit.mean$residuals,1)^2) +b1 * tail(fit.var$fit[,1],1)

# Finally, we compute the 95% forecast interval:
f1 + -1 * 1.96 * sqrt(h1)
f1 + 1 * 1.96 * sqrt(h1)
```

### Figure 11. Conditional Variances
Here are the conditional variances, with the last 10 values printed out:

```{r}
ht <- fit.var$fit[,1]^2
tail(ht, n=10)
```


Here is a plot of the conditional variances:

```{r}
# Add plot of the conditional variances
plot(date, ht, type="l", col=4)
```

```{r echo=FALSE}
# Exporting the conditional variances table for website:
condition.var.table <- data.frame(date, ht)
write.csv(condition.var.table, file="/Users/chuamelia/Desktop/ameliachu.github.io/realtime_bike/condition-var.csv", row.names = FALSE)
```

Here is a time series plot which simultaneously shows the log bikes available,
together with the ARIMA-ARCH one-step-ahead 95% forecast intervals based on
information available in the previous day:

```{r}
upper <- f + 1.96 * sqrt(ht)
lower <- f - 1.96 * sqrt(ht)
log.status.intervals <- paste(upper,";", log.status,";", lower,sep="")
forecast.int <- data.frame(date, log.status.intervals)
write.csv(forecast.int, file="/Users/chuamelia/Desktop/ameliachu.github.io/realtime_bike/12b_forecast-int.csv", row.names = FALSE)
```

```{r fig.width=12, fig.height=8}
plot(date, log.status, type="l")
lines(date, f + 1.96 * sqrt(ht), lty=2, col=2)
lines(date,  f - 1.96 * sqrt(ht), lty=2, col=2)
```

```{r}
plot(forecast(fit.mean, h=100, level=95), col=2)
arima.interval <- data.frame(forecast(fit.mean, h=100, level=95))
log.status.actuals <- c(paste(log.status,";", log.status,";",log.status,sep=""), rep("null;null;null", nrow(arima.interval)))
log.status.forecasts <- c(rep("null;null;null", length(log.status)), paste(arima.interval$Lo.95,";", arima.interval$Point.Forecast,";", arima.interval$Hi.95,sep=""))
time <- 1:length(log.status.actuals)
arima.forecasts <- data.frame(time, log.status.actuals, log.status.forecasts)
write.csv(arima.forecasts, file="/Users/chuamelia/Desktop/ameliachu.github.io/realtime_bike/12a_arima-forecast-int.csv", row.names = FALSE)
```

Here is a normal probability plot of the ARCH residuals.

```{r}
library("e1071")  
# Here we compute the arch residuals:

resid.arch <- resid / sqrt(ht)
plot(date, resid.arch, col=4, type="l")

#Here is the normal probability plot:
qqnorm(resid.arch)
qqlinetable <- data.frame(qqline(resid.arch))

kurtosis(resid.arch, na.rm=TRUE)
#qq.resid.arch <- data.frame(qqnorm(resid.arch, plot.it = FALSE))
#write.csv(qq.resid.arch, file="/Users/chuamelia/Desktop/ameliachu.github.io/realtime_bike/13_qqplot.csv", row.names = FALSE)
```


```{r}
#https://stackoverflow.com/questions/37897274/plot-lm-extracting-numbers-labelled-in-the-diagnostic-q-q-plot
resid.arch <- na.omit(resid.arch)
x <- sort(abs(resid.arch), decreasing = TRUE)
x[1:4]
```


Here is a count of how many prediction interval failures there were:

```{r}
# Count the number of times the prediciton
# interval failed:

sum(abs(resid.arch) >  1.96, na.rm=TRUE)
```

The number of prediction intervals is:
```{r}
sum(!is.na(resid.arch))
```
